Predicting Severity of _C. difficile_ Infections from the Taxonomic Composition of the Gut Microbiome

### 2023-06-07 paper/paper.qmd
More text to introduce AUBPRC in results

We first set out to train the best models possible for each severity definition.
Not all samples have outcomes available for all four severity definitions due to
missing data for some patient lab values and incomplete chart review
(@fig-flowchart B), thus each severity definition had a different number of
samples when using as many samples as possible (@tbl-counts A).
We referred to these as the full datasets.
Random forest models were trained on 100 splits of the datasets into training
and test sets, and performance was evaluated on the held-out test set using the
area under the receiver-operator characteristic curve (AUROC).
Since the severity classes were highly imbalanced with different proportions of
severe samples between definitions, we also calculated the balanced precision
and the area under the balanced precision-recall curve (AUBPRC) as first
proposed by @wu_improved_2021 to describe the precision that would be expected
if the outcome classes were balanced.

@@

The test set median AUBPRCs from the full datasets followed a similar pattern as 
the test set AUROCs with
`r perf_medians$aubprc_full_idsa` for IDSA severity,
`r perf_medians$aubprc_full_allcause` for all-cause severity,
`r perf_medians$aubprc_full_attrib` for attributable severity, and
`r perf_medians$aubprc_full_pragmatic` for pragmatic severity.
The AUBPRCs were significantly different from each other (P < 0.05) for
each pair of severity definitions except for attributable vs all-cause.
We plotted the balanced precision-recall curve and found that the IDSA definition
outperformed all other models at very low recall values, but the others outperform
IDSA at all other points of the curve (@fig-performance C).
The 95% confidence intervals overlapped the baseline AUROC and AUBPRC for
the attributable severity models, while all others did not overlap the baseline.

@@

The workflow was defined with Snakemake [@koster_snakemake_2012] and
dependencies were managed with conda environments.
Scripts were written in R [@r_core_team_r_2020],
Python [@van_rossum_python_2009],
and GNU bash.
Additional software and packages used in the creation of this manuscript include
cowplot [@wilke_cowplot_2020],
ggtext [@wilke_ggtext_2020],
ggsankey [@sjoberg_ggsankey_2022],
schtools [@sovacool_schtools_2022],
the tidyverse metapackage [@wickham_welcome_2019],
Quarto, and
vegan [@oksanen_vegan_2023].

>>>>

We first set out to train the best models possible for each severity definition.
Not all samples have outcomes available for all four severity definitions due to
missing data for some patient lab values and incomplete chart review
(@fig-flowchart B), thus each severity definition had a different number of
samples when using as many samples as possible (@tbl-counts A).
We referred to these as the full datasets.
Random forest models were trained on 100 splits of the datasets into training
and test sets, and performance was evaluated on the held-out test set using the
area under the receiver-operator characteristic curve (AUROC).
Since the severity outcomes were highly imbalanced with different proportions of
severe samples between definitions, we also calculated the balanced precision
and the area under the balanced precision-recall curve (AUBPRC) as first
proposed by @wu_improved_2021 to describe the precision that would be expected
if the outcomes were balanced.

@@

Since the data are highly imbalanced with only a small proportion of CDI cases
having a severe outcome, evaluating the trade-off between precision and recall
is more informative than the receiver-operator characteristic because precision
and recall do not consider true negatives, which may overinflate the AUROC.
However, unlike for AUROC, the baseline for the area under the precision-recall
curve depends on the proportion of positive outcomes (i.e. severe cases) in the
data, which vary across these severity definitions.
To allow comparison of precision across datasets with different proportions of
positives, @wu_improved_2021 introduced the concept of balanced precision, a
transformation of precision based on Bayes' theorem that represents the
precision that would have been expected if the proportion of positives were
balanced at 0.5.
Reporting the area under the balanced precision-recall curve (AUBPRC) allows us
to compare the trade-off between precision and recall for our different severity
defintions.
The test set median AUBPRCs from the full datasets followed a similar pattern as 
the test set AUROCs with
`r perf_medians$aubprc_full_idsa` for IDSA severity,
`r perf_medians$aubprc_full_allcause` for all-cause severity,
`r perf_medians$aubprc_full_attrib` for attributable severity, and
`r perf_medians$aubprc_full_pragmatic` for pragmatic severity.
The AUBPRCs were significantly different from each other (P < 0.05) for
each pair of severity definitions except for attributable vs all-cause.
We plotted the balanced precision-recall curve and found that the IDSA definition
outperformed all other models at very low recall values, but the others outperform
IDSA at all other points of the curve (@fig-performance C).
The 95% confidence intervals overlapped the baseline AUROC and AUBPRC for
the attributable severity models, while all others did not overlap the baseline.

@@

The workflow was defined with Snakemake [@koster_snakemake_2012] and
dependencies were managed with conda environments.
Scripts were written in R [@r_core_team_r_2023],
Python [@van_rossum_python_2009],
and GNU bash.
Additional software and packages used in the creation of this manuscript include
cowplot [@wilke_cowplot_2020],
ggtext [@wilke_ggtext_2020],
ggsankey [@sjoberg_ggsankey_2022],
schtools [@sovacool_schtools_2022],
the tidyverse metapackage [@wickham_welcome_2019],
Quarto, and
vegan [@oksanen_vegan_2023].
